---
title: "Public perception of marine habitats in the GBR - preliminary results"
author: "Riccardo Losciale"
date: "03/08/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo= FALSE}
pp_df <- read.csv("../data/public_perception_data/public_perception.csv")

```
## Sample demographic
```{r echo = FALSE}
library("ggplot2")
library("dplyr")
age_distribution <- ggplot(data=subset(pp_df, !is.na(Age)), aes(x=Age)) +
  geom_bar(mapping = aes(Age,stat(prop), group = 1)) +
  theme_classic() +
  coord_flip()

gender_distribution <- ggplot (data = pp_df) +
  geom_bar(mapping = aes(x = Gender, y = stat(prop), group = 1)) +
             scale_y_continuous(limits = c(0,0.75))

pp_df <- within(pp_df, 
                   Education <- factor(Education, 
                                      levels=names(sort(table(Education), 
                                                        decreasing = TRUE))))
ed_plot <- ggplot (data = pp_df) +
  geom_bar(mapping = aes(x = Education, y = stat(prop), group = 1)) +
             coord_flip()


pp_df <- within(pp_df, 
                main_activity_trim <- factor(main_activity_trim, 
                                    levels=names(sort(table(main_activity_trim), 
                                                      decreasing=F))))
activities_plot <- ggplot(data =subset(pp_df, !is.na(main_activity_trim)), aes(x=main_activity_trim)) +
  geom_bar(mapping = aes(x = main_activity_trim, y = stat(prop), group = 1)) +
             coord_flip()

pp_df <- within(pp_df, 
                visit_freq <- factor(visit_freq, 
                                        levels=names(sort(table(visit_freq), 
                                                          decreasing=F))))
visit_plot <- ggplot(data = subset(pp_df, !is.na(visit_freq))) +
  geom_bar(mapping = aes(x = visit_freq, y = stat(prop), group = 1)) +
             coord_flip()
recreation <- tribble( 
~activity, ~count,
"Scuba", 41,
"Recreationa fishing", 50,
"Spearfishing", 20,
"Sailing", 46, 
"Snorkelling / Freediving", 64,
"Boating", 36,
"Jetski", 8,
"Kite surfing", 8,
"Kayak", 2)
rec_pl <- recreation %>% ggplot(aes(x = reorder(activity, count), y = count)) + geom_bar(stat = "identity") + coord_flip()

#install.packages("ggpubr")
library("ggpubr")
demographic <- ggarrange(age_distribution, gender_distribution, ed_plot, activities_plot, visit_plot, rec_pl, ncol = 2, nrow = 3, labels = c("A", "B", "C", "D", "E", "F"))
demographic
pp_df %>% 
  count(Age)
pp_df %>%
  count(Gender)
pp_df %>%
  count(Education)
pp_df %>%
  count(main_activity)
pp_df %>%
  count(visit_freq)
recreation %>%
  arrange(desc(count))
```

## Change factors levels in dataframe
```{r echo= T}
library("forcats")
Age2 <- fct_collapse(pp_df$Age,
                         young_adults = c("18-24", "25-34"), 
                     middle_aged = c("35-44", "45-54"),
                         older_adults = c("55-64", "65+"))
Education2 <- fct_collapse(pp_df$Education,
                           secondary_education = c("Secondary Education", "Other", "Technical College "),
                           higher_education = c("Bachelor's degree", "Diploma", "Master's degree", "Ph. D"))
seagrass_perception <- fct_collapse(pp_df$seagrass_image,
                               healthy = c("good","very_good"),
                               unhealthy = c("poor","very_poor"))
coral_perception <- fct_collapse(pp_df$coral_image,
                               healthy = c("good","very_good"),
                               unhealthy = c("poor","very_poor"))
activity_type <- fct_collapse(pp_df$main_activity_trim,
                              above_water = c("Boating", "Jetski","Kayak", "Kite_surfing", "Rec_fish", "Sailing"),
                              below_water = c("Scuba", "Snorkeling_freediving", "Spearfishing"))

visit_freq2 <- fct_collapse(pp_df$visit_freq, 
                            frequent_visitor = c("Every day","About once a week", "Once a month"), 
                            non_frequent_visitor = c("Few times a year", "Once a year", "Less than once a year"))
# Add new column to dataframe with cbind
pp_df <- cbind(pp_df, Age2, Education2, seagrass_perception, coral_perception, activity_type, visit_freq2)
age2_plot <- ggplot(data =subset(pp_df, !is.na(Age2)), aes(x=Age2)) +
  geom_bar(mapping = aes(x = Age2, y = stat(prop), group = 1)) +
             coord_flip()
edu2_plot <- ggplot(data =subset(pp_df, !is.na(Education2)), aes(x=Education2)) +
  geom_bar(mapping = aes(x = Education2, y = stat(prop), group = 1)) +
             coord_flip()
act2_plot <- ggplot(data =subset(pp_df, !is.na(activity_type)), aes(x=activity_type)) +
  geom_bar(mapping = aes(x = activity_type, y = stat(prop), group = 1)) +
             coord_flip()
visit2_plot <- ggplot(data =subset(pp_df, !is.na(visit_freq2)), aes(x=visit_freq2)) +
  geom_bar(mapping = aes(x = visit_freq2, y = stat(prop), group = 1)) +
             coord_flip()
demographic2 <- ggarrange(age2_plot, edu2_plot, act2_plot, visit2_plot, ncol = 2, nrow = 2, labels = c("a","b","c","d"))
demographic2
                
```

## Image selection responses
```{r echo = F}
pp_df$seagrass_image <- factor(pp_df$seagrass_image,levels = c("very_poor", "poor", "good", "very_good"))
pl1 <- pp_df %>%
  filter(!is.na(seagrass_image)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = seagrass_image, y = stat(prop), group = 1 ))

pl2 <- pp_df %>%
  filter(!is.na(seagrass_perception)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = seagrass_perception, y = stat(prop), group = 1 ))
pp_df$coral_image <- factor(pp_df$coral_image,levels = c("very_poor", "poor", "good", "very_good"))
pl3 <- pp_df %>%
  filter(!is.na(coral_image)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = coral_image, y = stat(prop), group = 1 ))

pl4 <- pp_df %>%
  filter(!is.na(coral_perception)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = coral_perception, y = stat(prop), group = 1 ))
perception <- ggarrange(pl1, pl2, pl3, pl4, ncol = 2, nrow = 2, labels = c("A","B","C","D"))
perception

```


Based on image selection, we can see that more than 80% of marine recreational users (MCUs) have the perception that coral reefs are in good or very good conditions. 
However, 60% of MCUs have the perception that seagrass are in poor or very poor conditions. 

Probability of the outcome  good/very good coral image choiche :3.74063e-13
Probability of the outcome  poor/very poor seagrass image choiche: 0.01116688

## Perception of seagrass status might be correlated with Age and activity type
```{r echo = F}
pl5 <- pp_df %>% 
  filter(!is.na(seagrass_perception) & !is.na(Age2)) %>%
  count(Age2, seagrass_perception) %>%  
  ggplot() +
    geom_tile(mapping = aes(x = Age2, y = seagrass_perception, fill = n))
pl6 <- pp_df %>% 
  filter(!is.na(seagrass_perception) & !is.na(activity_type)) %>%
  count(activity_type, seagrass_perception) %>%  
  ggplot() +
    geom_tile(mapping = aes(x = activity_type, y = seagrass_perception, fill = n))
correlation <- ggarrange(pl5, pl6, ncol = 1, nrow = 2, labels = c("A","B"))
correlation
library("corrplot")
age_sea <- pp_df %>%
  select(Age2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Age2))
age_sea <- table(age_sea$Age2, age_sea$seagrass_perception)
act_sea1 <- pp_df %>%
  select(activity_type, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(activity_type))
act_sea1 <- table(act_sea1$activity_type, act_sea1$seagrass_perception)
age_act <- pp_df %>%
  select(Age2, activity_type) %>%
  filter(!is.na(Age2) & !is.na(activity_type))
age_act <- table(age_act$Age2, age_act$activity_type)

chis_age <- chisq.test(age_sea)
chis_activity1 <- chisq.test(act_sea1)

corrplot(chis_age$residuals, is.cor = FALSE)
corrplot(chis_activity1$residuals, is.cor = FALSE)
prop.table(age_sea, 1) 
prop.table(age_sea, 2) 
chis_age
prop.table(act_sea1, 1) 
prop.table(act_sea1, 2) 

chis_activity1
mosaicplot(age_act, shade = TRUE, las=2,
           main = "Perception of seagrass status")
prop.table(age_act, 1) # row percentages
prop.table(age_act, 2) # column percentages
chis_age_act <- chisq.test(age_act)
chis_age_act

```
Chi square test of indipendence shows that the variables age and activity type and perception of seagrass status are not indipendent (p < 0.05). 76% of people between 18 and 35 years old selected the poor or very poor picture, whereas 65% of people older than 55 selected the good or very good picture. Similarly, 87% of people performing below water activities selected the poor/very poor image. However, 58.7% of people performing below water activities are younger adults.


## Seagrass and coral perception are not correlated
```{r echo = F}
library("gplots")
library("vcd")
sea_cor <- pp_df %>%
  select(seagrass_perception, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(seagrass_perception))
sea_cor <- table(sea_cor$seagrass_perception, sea_cor$coral_perception)
balloonplot(t(sea_cor), main = "Perception of seagrass status against coral status", xlab ="",
           ylab = "", label = F, show.margins = F)
assoc(head(sea_cor, 5), shade = TRUE, las=3)
edu_sea <- pp_df %>%
  select(Education, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Education))
edu_sea <- table(edu_sea$Education, edu_sea$seagrass_perception)
edu_sea1 <- pp_df %>%
  select(Education2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Education2))
edu_sea1 <- table(edu_sea1$Education2, edu_sea1$seagrass_perception)
chis_edu <- chisq.test(edu_sea)
chis_edu1 <- chisq.test(edu_sea1)
chis_edu
chis_edu1
gender_sea <- table(pp_df$Gender, pp_df$seagrass_perception)
balloonplot(t(gender_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
visit_sea <- pp_df %>%
  select(visit_freq, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(visit_freq))
visit_sea <- table(visit_sea$visit_freq, visit_sea$seagrass_perception)
balloonplot(t(visit_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
visit_sea1 <- pp_df %>%
  select(visit_freq2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(visit_freq2))
visit_sea1 <- table(visit_sea1$visit_freq2, visit_sea1$seagrass_perception)
balloonplot(t(visit_sea1), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
chis_visit1 <- chisq.test(visit_sea1)
chis_visit1
```

Education level, gender and visit frequency are not good predictor of perception of seagrass status.
Additionally, selection of seagrass image and coral image are not correlated
Based on Chi square statistic, there is no correlation between coral perception and any of the dependent variables.
