---
title: "Public perception of marine habitats in the GBR"
author: "Riccardo Losciale"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Open csv file
```{r echo= FALSE}
pp_df <- read.csv("../data/public_perception_data/public_perception.csv")

```

# Explore dependent variables / demographics
```{r echo =FALSE}
library("ggplot2")
library("dplyr")
age_distribution <- ggplot(data=subset(pp_df, !is.na(Age)), aes(x=Age)) +
  geom_bar(mapping = aes(Age,stat(prop), group = 1)) +
  theme_classic() +
  coord_flip()

gender_distribution <- ggplot (data = pp_df) +
  geom_bar(mapping = aes(x = Gender, y = stat(prop), group = 1)) +
             scale_y_continuous(limits = c(0,0.75))

pp_df <- within(pp_df, 
                   Education <- factor(Education, 
                                      levels=names(sort(table(Education), 
                                                        decreasing = TRUE))))
ed_plot <- ggplot (data = pp_df) +
  geom_bar(mapping = aes(x = Education, y = stat(prop), group = 1)) +
             coord_flip()


pp_df <- within(pp_df, 
                main_activity_trim <- factor(main_activity_trim, 
                                    levels=names(sort(table(main_activity_trim), 
                                                      decreasing=F))))
activities_plot <- ggplot(data =subset(pp_df, !is.na(main_activity_trim)), aes(x=main_activity_trim)) +
  geom_bar(mapping = aes(x = main_activity_trim, y = stat(prop), group = 1)) +
             coord_flip()

pp_df <- within(pp_df, 
                visit_freq <- factor(visit_freq, 
                                        levels=names(sort(table(visit_freq), 
                                                          decreasing=F))))
visit_plot <- ggplot(data = subset(pp_df, !is.na(visit_freq))) +
  geom_bar(mapping = aes(x = visit_freq, y = stat(prop), group = 1)) +
             coord_flip()

#install.packages("ggpubr")
library("ggpubr")
demographic <- ggarrange(age_distribution, gender_distribution, ed_plot, activities_plot, visit_plot, ncol = 3, nrow = 2, labels = c("A", "B", "C", "D", "E"))

demographic
pp_df %>%
  count(Age)
pp_df %>%
  count(Gender)
pp_df %>%
  count(Education)
pp_df %>%
  count(main_activity)
pp_df %>%
  count(visit_freq)
```


## Change factors levels in dataframe
```{r echo= T}
library("forcats")
Age2 <- fct_collapse(pp_df$Age,
                         young_adults = c("18-24", "25-34"), 
                     middle_aged = c("35-44", "45-54"),
                         older_adults = c("55-64", "65+"))
Education2 <- fct_collapse(pp_df$Education,
                           secondary_education = c("Secondary Education", "Other", "Technical College "),
                           higher_education = c("Bachelor's degree", "Diploma", "Master's degree", "Ph. D"))
seagrass_perception <- fct_collapse(pp_df$seagrass_image,
                               healthy = c("good","very_good"),
                               unhealthy = c("poor","very_poor"))
coral_perception <- fct_collapse(pp_df$coral_image,
                               healthy = c("good","very_good"),
                               unhealthy = c("poor","very_poor"))
activity_type <- fct_collapse(pp_df$main_activity_trim,
                              above_water = c("Boating", "Jetski","Kayak", "Kite_surfing", "Rec_fish", "Sailing"),
                              below_water = c("Scuba", "Snorkeling_freediving", "Spearfishing"))

visit_freq2 <- fct_collapse(pp_df$visit_freq, 
                            frequent_visitor = c("Every day","About once a week", "Once a month"), 
                            non_frequent_visitor = c("Few times a year", "Once a year", "Less than once a year"))
# Add new column to dataframe with cbind
pp_df <- cbind(pp_df, Age2, Education2, seagrass_perception, coral_perception, activity_type, visit_freq2)
                
```

## Correlation between Age and perception of seagrass status
```{r echo = FALSE}
pp_df$seagrass_image <- factor(pp_df$seagrass_image,levels = c("very_poor", "poor", "good", "very_good"))
pp_df %>%
  filter(!is.na(seagrass_image)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = seagrass_image, y = stat(prop), group = 1 ))

pp_df %>%
  filter(!is.na(seagrass_perception)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = seagrass_perception, y = stat(prop), group = 1 ))

pp_df %>% 
  filter(!is.na(seagrass_perception) & !is.na(Age2)) %>%
  count(Age2, seagrass_perception) %>%  
  ggplot() +
    geom_tile(mapping = aes(x = Age2, y = seagrass_perception, fill = n))

```



## Correlation between Age and perception of seagrass status
```{r echo= FALSE}
age_sea <- pp_df %>%
  select(Age2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Age2))
age_sea <- table(age_sea$Age2, age_sea$seagrass_perception)
library("gplots")
balloonplot(t(age_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
library("latticeExtra")
library("vcd")
assoc(head(age_sea, 5), shade = TRUE, las=3)
prop.table(age_sea) # cell percentages
prop.table(age_sea, 1) # row percentages
prop.table(age_sea, 2) # column percentages
chis_age <- chisq.test(age_sea)
chis_age
chis_age$observed
round(chis_age$expected,2)
round(chis_age$residuals, 3)
library(corrplot)
corrplot(chis_age$residuals, is.cor = FALSE)
# Contibution in percentage (%)
contrib <- 100*chis_age$residuals^2/chis_age$statistic
round(contrib, 3)
# Visualize the contribution
corrplot(contrib, is.cor = FALSE)
```

Chi square test of indipendence shows that the variables age and perception of seagrass status are not indipendent (p < 0.05). These graphs show that people between 18 and 35 years old selected more often the very poor or poor image of seagrass compared to people older than 55 years, who maily selected the good or very good image.

## Correlation between Education level and image choiche of seagrass  levels
```{r echo = FALSE}
edu_sea <- pp_df %>%
  select(Education, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Education))
edu_sea <- table(edu_sea$Education, edu_sea$seagrass_perception)
edu_sea
balloonplot(t(edu_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
chis_edu <- chisq.test(edu_sea)
chis_edu
edu_sea1 <- pp_df %>%
  select(Education2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(Education2))
edu_sea1 <- table(edu_sea1$Education2, edu_sea1$seagrass_perception)

balloonplot(t(edu_sea1), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
mosaicplot(edu_sea1, shade = TRUE, las=2,
           main = "Perception of seagrass status")

prop.table(edu_sea1) # cell percentages
prop.table(edu_sea1, 1) # row percentages
prop.table(edu_sea1, 2) # column percentages
chis_edu1 <- chisq.test(edu_sea1)
chis_edu1
chis_edu1$observed
round(chis_edu1$expected,2)
round(chis_edu1$residuals, 3)
corrplot(chis_edu1$residuals, is.cor = FALSE)
# Contibution in percentage (%)
contrib_edu1 <- 100*chis_edu1$residuals^2/chis_edu1$statistic
round(contrib_edu1, 3)
# Visualize the contribution
corrplot(contrib_edu1, is.cor = FALSE)

```
This results shows that there is not enough confidence to state that the variables education levels and seagrass perception are not indipendent. However, unhealthy seagrass pictures were selected by 67% of respondents with Higher education. 

## Correlation between gender and image choiche of seagrass with new levels
```{r echo = FALSE}
gender_sea <- table(pp_df$Gender, pp_df$seagrass_perception)
gender_sea
balloonplot(t(gender_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)

```
No difference between gender


## Correlation between main activity and image choiche of seagrass with original levels
```{r echo = F}
act_sea <- pp_df %>%
  select(main_activity_trim, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(main_activity_trim))
act_sea <- table(act_sea$main_activity_trim, act_sea$seagrass_perception)

act_sea
balloonplot(t(act_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
act_sea1 <- pp_df %>%
  select(activity_type, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(activity_type))
act_sea1 <- table(act_sea1$activity_type, act_sea1$seagrass_perception)

act_sea1
balloonplot(t(act_sea1), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
mosaicplot(act_sea1, shade = TRUE, las=2,
           main = "Perception of seagrass status")
assoc(head(act_sea1, 5), shade = TRUE, las=3)
prop.table(act_sea1) # cell percentages
prop.table(act_sea1, 1) # row percentages
prop.table(act_sea1, 2) # column percentages
chis_activity1 <- chisq.test(act_sea1)
chis_activity1
chis_activity1$observed
round(chis_activity1$expected,2)
round(chis_activity1$residuals, 3)
corrplot(chis_activity1$residuals, is.cor = FALSE)
# Contibution in percentage (%)
contrib_activity1 <- 100*chis_activity1$residuals^2/chis_activity1$statistic
round(contrib_activity1, 3)
# Visualize the contribution
corrplot(contrib_activity1, is.cor = FALSE)

```
Chi square test of indipendece shows that the variables activity type and seagrass image choiche are statistically not indipendent.  86.8% of people performing below water activities have selected an unhealthy seagrass picture. On the other hand, 87.5% of people chosing a helathy seagrass picture perform activities above water. 

## Correlation between visit frequency and image choiche of seagrass with original levels
```{r echo = F}
visit_sea <- pp_df %>%
  select(visit_freq, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(visit_freq))
visit_sea <- table(visit_sea$visit_freq, visit_sea$seagrass_perception)
visit_sea
balloonplot(t(visit_sea), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
visit_sea1 <- pp_df %>%
  select(visit_freq2, seagrass_perception) %>%
  filter(!is.na(seagrass_perception) & !is.na(visit_freq2))
visit_sea1 <- table(visit_sea1$visit_freq2, visit_sea1$seagrass_perception)
visit_sea1

balloonplot(t(visit_sea1), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
mosaicplot(visit_sea1, shade = TRUE, las=2,
           main = "Perception of seagrass status")
assoc(head(visit_sea1, 5), shade = TRUE, las=3)

chis_visit1 <- chisq.test(visit_sea1)
chis_visit1

```
Results show no relation between visit Frequency and perception of seagrass status

## Multicollinearity among dependent variables
```{r echo = F}
pp_df %>% 
  filter(!is.na(Age2)) %>%
  count(Age2, Education2) %>%
  ggplot() +
  geom_tile(mapping = aes(x = Age2, y = Education2, fill = n))

age_edu <- pp_df %>%
  select(Age2, Education2) %>%
  filter(!is.na(Age2) & !is.na(Education2))
age_edu <- table(age_edu$Age2, age_edu$Education2)
age_edu
balloonplot(t(age_edu), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
assoc(head(age_edu, 5), shade = TRUE, las=3)

chis_age_edu <- chisq.test(age_edu)
chis_age_edu

pp_df %>% 
   filter(!is.na(Age2) & !is.na(activity_type)) %>% 
  count(Age2, activity_type) %>%
  ggplot() +
  geom_tile(mapping = aes(x = Age2, y = activity_type, fill = n))
age_act <- pp_df %>%
  select(Age2, activity_type) %>%
  filter(!is.na(Age2) & !is.na(activity_type))
age_act <- table(age_act$Age2, age_act$activity_type)
age_act

mosaicplot(age_act, shade = TRUE, las=2,
           main = "Perception of seagrass status")
assoc(head(age_act, 5), shade = TRUE, las=3)
prop.table(age_act) # cell percentages
prop.table(age_act, 1) # row percentages
prop.table(age_act, 2) # column percentages
chis_age_act <- chisq.test(age_act)
chis_age_act
chis_age_act$observed
round(chis_age_act$expected,2)
round(chis_age_act$residuals, 3)
corrplot(chis_age_act$residuals, is.cor = FALSE)
contrib_age_act <- 100*chis_age_act$residuals^2/chis_age_act$statistic
round(contrib_age_act, 3)
corrplot(contrib_age_act, is.cor = FALSE)

pp_df %>% 
  filter(!is.na(Age2), !is.na(visit_freq2)) %>%
  count(Age2, visit_freq2) %>% 
  ggplot() +
  geom_tile(mapping = aes(x = Age2, y = visit_freq2, fill = n))
age_freq <- pp_df %>%
  select(Age2, visit_freq2) %>%
  filter(!is.na(Age2) & !is.na(visit_freq2))
age_freq <- table(age_freq$Age2, age_freq$visit_freq2)
age_freq

balloonplot(t(age_freq), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
assoc(head(age_freq, 5), shade = TRUE, las=3)

```
No collinearity between: Age and Education, Age and visit frequency.  However there is multicollinearity between age and activity type. Youger people are associated with below water activities (scuba, snorkelling, spearfishing), whereas adults are more related to above water activities


## Correlation between demographic variables and perception of coral status
```{r echo=FALSE}
pp_df$coral_image <- factor(pp_df$coral_image,levels = c("very_poor", "poor", "good", "very_good"))
pp_df %>%
  filter(!is.na(coral_image)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = coral_image, y = stat(prop), group = 1 ))

pp_df %>%
  filter(!is.na(coral_perception)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = coral_perception, y = stat(prop), group = 1 ))

pp_df %>% 
  filter(!is.na(coral_perception) & !is.na(Age2)) %>%
  count(Age2, coral_perception) %>%  
  ggplot() +
    geom_tile(mapping = aes(x = Age2, y = coral_perception, fill = n))
```

Results shows that 84% of marine recreational users have a good or very good perception of the staus of coral reef 



## Correlation between Education and image choiche of corals with new levels
```{r echo=FALSE}
edu_cor <- pp_df %>%
  select(Education2, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(Education2))
edu_cor <- table(edu_cor$Education2, edu_cor$coral_perception)

balloonplot(t(edu_cor), main = "Perception of coral status", xlab ="",
           ylab = "", label = F, show.margins = F)
assoc(head(edu_cor, 5), shade = TRUE, las=3)
```
No correlation between age and image choice 


## Correlation between main activity and image choiche of corals with new levels
```{r echo = F}
activity_cor <- pp_df %>%
  select(activity_type, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(activity_type))
activity_cor <- table(activity_cor$activity_type, activity_cor$coral_perception)

balloonplot(t(activity_cor), main = "Perception of coral status", xlab ="",
           ylab = "", label = F, show.margins = F)
mosaicplot(activity_cor, shade = TRUE, las=2,
           main = "Perception of coral status")
assoc(head(activity_cor, 5), shade = TRUE, las=3)
```

No correlation between activity type and coral image choice

## Correlation between visit frequency and image choiche of corals 
```{r echo = F}
visit_cor <- pp_df %>%
  select(visit_freq, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(visit_freq))
visit_cor <- table(visit_cor$visit_freq, visit_cor$coral_perception)
visit_cor
balloonplot(t(visit_cor), main = "Perception of coral status", xlab ="",
           ylab = "", label = F, show.margins = F)
assoc(head(visit_cor, 5), shade = TRUE, las=3)

visit_cor1 <- pp_df %>%
  select(visit_freq2, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(visit_freq2))
visit_cor1 <- table(visit_cor1$visit_freq2, visit_cor1$coral_perception)
visit_cor1


balloonplot(t(visit_cor1), main = "Perception of seagrass status", xlab ="",
           ylab = "", label = F, show.margins = F)
mosaicplot(visit_cor1, shade = TRUE, las=2,
           main = "Perception of seagrass status")
assoc(head(visit_cor1, 5), shade = TRUE, las=3)

```

No correlation between visit frequency and coral image

## Multicollinearity between seagrass image choiche and coral image choiche
```{r echo = F}
sea_cor <- pp_df %>%
  select(seagrass_perception, coral_perception) %>%
  filter(!is.na(coral_perception) & !is.na(seagrass_perception))
sea_cor <- table(sea_cor$seagrass_perception, sea_cor$coral_perception)
sea_cor
balloonplot(t(sea_cor), main = "Perception of seagrass status against coral status", xlab ="",
           ylab = "", label = F, show.margins = F)

assoc(head(sea_cor, 5), shade = TRUE, las=3)

```
No multicollinearity between image choiche of seagrass and corals

## Climate change belif 
```{recho=F}
(pp_df %>%
  filter(!is.na(cc_bel), !is.na(Age2)) %>%
  ggplot(mapping = aes(x = cc_bel, color = Age2)) + 
  geom_freqpoly(bins = 10) +
  scale_x_continuous(limits = c(0,100)))

                

```


## Analysis of benefits
```{r}
benefits_df <- pp_df %>% 
  select(Age, Age2, Education, Education2, visit_freq, visit_freq2, Gender, main_activity_trim, activity_type, ends_with("_ben") ) 

benefits_only <- pp_df %>%
  select(ends_with("_ben")) 
boxplot(benefits_only)
summary(benefits_only)
(benefits_df %>%
  filter(!is.na(Age2) & !is.na(recr_fish_ben) & !is.na(activity_type)) %>%
  ggplot() +
  geom_boxplot(mapping = aes(activity_type, recr_fish_ben)) +
  coord_flip())
```
 
